---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

# Packages 
```{r Packages, message=FALSE}
library(tidyverse)
```
# Loading and preprocessing the data
The data set, called __*activity.csv*__, comes in a *Winrar* file, I extracted it and the data was saved in a new file called __*activity*__.

### Raw data

```{r Raw data, cache=TRUE}
Walking <- 
  read.csv(file = 'activity/activity.csv')

head(Walking, 3)
```

### Summary by Day
Ignoring the missing values, I created two data sets: 

  * **Day_step**, where groups were made based on the date.
  
  * **Interval_step**, where groups were made based on the 5-minutes intervals.


```{r Summaries, cache=TRUE}
Day_step <-  Walking %>% 
               group_by(date) %>% 
                  summarise(Total = sum(steps, na.rm = TRUE))

head(Day_step,3)

Interval_step <-  Walking %>% 
                    group_by(interval) %>% 
                      summarise(Mean = mean(steps, na.rm = TRUE))
head(Interval_step,3)
```
## What is mean total number of steps taken per day?
```{r Plot1}
Day_plot <- ggplot(data = Day_step)

Day_plot + 
  geom_histogram(aes(x = Total), color = 'black', binwidth = 2000) +
    xlab('Mean number of steps')

```

## What is the average daily activity pattern?
```{r Plot2}
ggplot(data = Interval_step) +
  geom_line(aes(x =interval, y = Mean)) +
    ylab('Mean number of steps') + xlab('Interval')

```
```{r echo=FALSE}
interval.max <- max(Interval_step$interval)

```
The interval with the **maximum** mean number of steps is: **`r interval.max`** 



## Imputing missing values
```{r Missing calue, echo=FALSE}
Missing <- sum(is.na(Walking$steps)) 

```
There are **`r Missing`** rows with missing values. I decided to fill the missing values with the median of that interval. 
```{r, cache=TRUE}
Day_step_No_NA <-  
  Walking %>%     
      group_by(interval) %>% 
          mutate(steps = ifelse(is.na(steps), median(steps, na.rm = TRUE),                     steps)) 

Day_step_No_NA
```

```{r Plot3}
Day_step_No_NA %>%
  group_by(date) %>% 
      summarise(Total = sum(steps)) %>%
          ggplot(aes(x = Total)) + 
              geom_histogram(color = 'black', binwidth = 2000) +
              xlab('Mean number of steps')

```
```{r echo=FALSE, message=FALSE}
day.mean2 <- mean(Day_step_No_NA$steps)
day.median2 <- median(Day_step_No_NA$steps)
day.mean <- mean(Day_step$steps, na.rm =TRUE)
day.median <- median(Day_step$steps, na.rm =TRUE)
```
  * The new **mean** number of steps is: **`r day.mean2`**
  * The new **median** number of steps is: **`r day.median2`**


## Are there differences in activity patterns between weekdays and weekends?

```{r }
Day_step_Week <-
  Day_step_No_NA %>%
    mutate(Weekday = ifelse(
      (weekdays(as.Date(date)) == "Saturday")|(weekdays(as.Date(date)) == "Sunday"), 'Weekend', 'Weekday'))

head(Day_step_Week, 3)
```
```{r message=FALSE}
Day_step_Week %>%
  group_by(interval, Weekday) %>%
    summarize(Mean = mean(steps)) %>%
        ggplot() +
            geom_line(aes(x =interval, y = Mean, color = Weekday)) +
            ylab('Mean number of steps') + xlab('Interval')
```
